---
title: "Billboard Graphics"
author: "Abdul Rahman"
execute:
  echo: false
---
```{r}
```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(tutorial.helpers)
  library(plotly)
})

```

```{r}
# Load the billboard dataset
billboard <- tidyr::billboard

# Convert the data from wide to long format for plotting
billboard_long <- billboard %>%
  pivot_longer(cols = starts_with("wk"), names_to = "week", values_to = "rank") %>%
  filter(!is.na(rank)) %>%
  mutate(week = as.numeric(str_remove(week, "wk")))

# Identify top 5 most frequently appearing artists
top_artists <- billboard_long %>%
  count(artist) %>%
  slice_max(n, n = 5) %>%
  pull(artist)

# Filter for top artists and tracks that reached Top 20
filtered_data <- billboard_long %>%
  filter(artist %in% top_artists, rank <= 20)

# Identify peak label positions for tracks
peak_labels <- filtered_data %>%
  group_by(track) %>%
  slice_min(rank, n = 1) %>%
  ungroup()

# Create refined plot
p <- filtered_data %>%
  ggplot(aes(x = week, y = rank, color = artist,
             group = track,
             text = paste0(
               "🎤 Artist: ", artist, "<br>",
               "🎶 Track: ", track, "<br>",
               "📅 Week: ", week, "<br>",
               "📈 Rank: ", rank))) +
  geom_line(alpha = 0.9, linewidth = 1.1) +
  geom_point(data = peak_labels, aes(x = week, y = rank), size = 2, color = "black") +
  scale_y_reverse(breaks = seq(1, 20, by = 2)) +
  facet_wrap(~ artist, scales = "free_y") +
  labs(
    title = "🎵 Billboard Hot 100 — Top Tracks from 5 Popular Artists",
    subtitle = "Showing performance of tracks that peaked within Top 20",
    x = "Weeks on Chart",
    y = "Billboard Rank (1 = Top)",
    color = "Artist"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 20),
    plot.subtitle = element_text(size = 14),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold")
  )

# Convert to interactive plot
interactive_plot <- ggplotly(p, tooltip = "text")
interactive_plot

# Save static version
ggsave("billboard_top_artists_plot.png", p, width = 12, height = 6)


```

